---
title: "Evaluating Evidence and Making Decisions using Bayesian Statistics"
subtitle: "ISCoP Conference 2021"  
author: "Mattan S. Ben-Shachar (@mattansb)"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
knit: (
  function(inputFile, encoding, ...) { 
    rmarkdown::render( 
      input       = inputFile, 
      encoding    = encoding,
      output_file = "index", 
      ...) })
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  # primary_color = "#1381B0",
  # secondary_color = "#FF961C",
  # inverse_header_color = "#FFFFFF"
)
```


Link to this presentation and materials on some GitHub repo.
This presentation is the "html page" for that repo.

---
class: inverse, center, middle

# What Is Bayes

---

Steal from new Gelman paper...

- Prior
- likelihood
- Posterior (is not sampling distribution)

We are talking about updating.

---
class: center, middle

## Why Bayes

---

## Why Bayes

- Complex models - use priors to restrict parameter search
- Talk about probabilities...


---
class: inverse, center, middle

# Making Decisions using Bayesian Statistics

---

Here we are limit our discussion to univariate parameters (slopes, variances) or estimates (contrasts, expected values).

We can also talk about multivariate desicion making, by comparing different models. See loo, BF, etc.

---

## Working example

Some Stroop or something?

- Working in `R`
- We can use `rstanarm` (conviniant) or `brms` (more flexable).
- We will use `bayestestR`

---
class: center, middle

## Posterior Based Measures

---


## Posterior Based Measures

- pd
- p-map
- p-rope


---
class: center, middle

## Likelihood Based Measures

---

## Likelihood Based Measures

- The Null-Interval Bayes Factor
- The Point-Null Bayes Factor
- Other Bayes Factors
  - Bayes factor for dividing hypotheses (link)
  - Restricted Model Bayes factors (link)
